True versus apparent shapes of bow shocks and bow waves

### 2017-11-03 papers/Paper1/sec-conic.tex
Port the conic figures to the new notation

@@

\newcommand\Sin{\ensuremath{\mathcal{S}}}
\newcommand\Cos{\ensuremath{\mathcal{C}}}
\newcommand\Cot{\ensuremath{\mathcal{T}}}

@@

%Instead of the excentricity, we utilize the parameter $\theta_c$ to characterize the different curves, where
%$\tan\theta_c = \frac{b}{a}$,  $b$ and $a$ are the typical parameters of conics. A positive value for $\theta_c$ indicates that the given curve is a closed one, i.e
%an ellipse, while a negative value indicates that is an hyperbola. %Insert figures if neccesary  
\begin{figure*}
\setlength\tabcolsep{2em}
\begin{tabular}{ll}
  (a) & (b) \\
  \includegraphics[height=0.45\linewidth]{figs/ellipse_edited}
      & \includegraphics[height=0.45\linewidth]{figs/hyperbola_edited}
\end{tabular}
\label{fig:conics}
\caption{Example off-center conic sections that can form quadrics of
  revolution: (a) ellipse, (b) hyperbola.  The relationship is shown
  between the conic section parameters \(a\), \(b\), \(\theta_c\) and the
  bowshock characteristic radii \(R_0\), \(R_{90}\), \(R_c\), as
  defined in Fig.~\ref{fig:characteristic-radii}. The origin (center
  of the weaker flow) is indicated by a red star, the center of
  curvature of the apex of the bow shock is indicated by a green dot,
  and the geometric center of the conic section is indicated by a blue
  cross.  }
\end{figure*}
\begin{figure}
\includegraphics[width=\linewidth]{conic1}
\caption{Example of a family of conic sections, all with the same
  normalized radius of curvature at their apex (marked by white dot):
  \(\widetilde{R}_c = R_c/R_0 = 2 \). The parameter
  $\theta_c = \pm \tan^{-1} b/a$ describes the family, with negative values
  corresponding to hyperbolae.  Differerent values of
  \(\widetilde{R}_c\) can be achieved by moving the origin along the
  \(x\)-axis while keeping the apex fixed at \(\tilde{x} = 1\).)}
\label{fig:conics-family}
\end{figure}

@@

A general parametric form\footnote{%
  The special case of the parabola needs to be treated differently,
  see Appendix~\ref{app:parabola}.} %
for the \(xy\) coordinates of the quadrics (in the \(\phi = 0\)plane, and
with the star at the origin) as a function of \(t = [0, \pi]\) is then
\begin{gather}
  \label{eq:par-xy}
  \begin{aligned}
    x &= x_0 + \sigma a \Cos(t) \\ 
    y &= b\Sin(t) 
  \end{aligned}
\end{gather}
where
\begin{equation}
  \label{eq:sin-sinh-etc}
  \Sin(t), \Cos(t) = \begin{cases}
    \sin(t), \cos(t) & \text{ellipse}\\
    \sinh(t), \cosh(t) & \text{hyperbola}
  \end{cases}
\end{equation}
Except for the circle case (\(\sigma = +1\), \(a = b\)), the parametric
variable \(t\) is not actually an angle in physical space.  Instead,
the polar form of the bow shape \(R(\theta)\) must be found by substituting
equations~\eqref{eq:par-xy} into \(\theta = \tan^{-1} y/x\) and
\(R = (x^2 + y^2)^{1/2}\).

The ``type'' of quadric surface can be characterized by the
\textit{quadric parameter}:
\begin{equation}
  \label{eq:Tq}
  T_q \equiv \sigma \frac{b} {a} \ , 
\end{equation}
where \(T_q < 0\) corresponds to open surfaces (hyperboloids) and
\(T_q > 0\) corresponds to closed surfaces (oblate spheroids with
\(T_q > 1\) and prolate spheroids with \(T_q < 1\)).  Special cases
are the sphere (\(T_q = 1\)) and the paraboloid (\(T_q = 0\)).  The
set of parameters \((T_q, a, x_0)\) are then sufficient to
characterize the star/bow combination, where \(a\) is the quadric
scale and \(x_0\) is its center displacement from the star.

But, we can also characterize the star/bow by \((R_0, \Pi, \Lambda)\), where
\(\Pi\) and \(\Lambda\) are the planitude and alatude, see
\S~\ref{sec:plan-alat-bow}.  We now derive the relationship between
these two descriptions.  The apex radius of curvature for a conic
section is
\begin{equation}
  \label{eq:Rc-conic}
  R_c = \frac{b^2}{a} \ ,
\end{equation}
whereas the perpendicular radius, \(R_{90}\), is the value of \(y\)
when \(x = 0\), which can be found from equations (\ref{eq:conic-x0},
\ref{eq:par-xy}) as
\begin{equation}
  \label{eq:conic-R90}
  R_{90} = \left|T_q\right|
  \left(2aR_0 - \sigma R_0^2\right)^{1/2} \ .
\end{equation}
Combining equations (\ref{eq:planitude}, \ref{eq:alatude},
\ref{eq:conic-x0}, \ref{eq:Tq}--\ref{eq:conic-R90}) yields
\begin{align}
  \label{eq:Tq-from-Pi-Lambda}
  T_q &= \sigma \left| 2 \Pi - \Lambda^2 \right|^{1/2} \\
  \label{eq:a-from-R0-Pi-Lambda}
  a &= R_0 \frac{ \Pi  } { \left|  2 \Pi - \Lambda^2\right| } \\
  x_0 &= R_0 \frac{\Pi - \Lambda^2} {2 \Pi - \Lambda^2}
\end{align}
From these we see that it is the sign of \(2 \Pi - \Lambda^2\) that determines
\(\sigma\) and thus whether a quadric is a spheroid or a hyperboloid.  For
example, for a constant planitude, \(\Pi\), we can have a family of
different quadric types, with varying alatude, \(\Lambda\), that increases
from oblate to prolate to hyperboloid, as illustrated in
Figure~\ref{fig:conics-family}.

@@

\begin{figure*}
  \centering
  \includegraphics[width=\linewidth]{figs/projected-R90-Rc-snapshots}
  \caption{Variation with inclination angle of projected quadric parameters}
  \label{fig:projected-R90-Rc-snapshots}
\end{figure*}
Once the parametrization is done, we can find the apparent shape of
the shell in the observer's frame, following the procedure explained
in section \ref{sec:projection}.  First of all, the intrinsic 3D shape of the shell is given by:
\begin{align}
x &= a\Cos(t)+x_0 \\ 
y &= b\Sin(t)\cos\phi \\
z &=  b\Sin(t)\sin\phi
\end{align}
The azimutal angle where the line of sight is tangent to the shell is given by equation (\ref{eq:tanphi}) and (\ref{eq:alpha}), then:
\begin{align}
\sin\phi_t &= \frac{b}{a}\tan i\Cot(t) 
\end{align}
where:
\begin{align}
\Cot(t) = \left\lbrace \begin{array}{c}
\cot t ~\mathrm{if~ellipse} \\
\coth t ~\mathrm{if~hyperbola}
\end{array}\right.
\end{align}

>>>>

\begin{figure*}
  \setlength\tabcolsep{2em}
  \begin{tabular}{ll}
    (a) & (b) \\
    \includegraphics[height=0.45\linewidth]{figs/ellipse_edited}
        & \includegraphics[height=0.45\linewidth]{figs/hyperbola_edited}
  \end{tabular}
  \caption{Example off-center conic sections that can form quadrics of
    revolution: (a) ellipse, (b) hyperbola.  The relationship is shown
    between the conic section parameters \(a\), \(b\), \(\theta_q\) and the
    bowshock characteristic radii \(R_0\), \(R_{90}\), \(R_c\), as
    defined in Fig.~\ref{fig:characteristic-radii}. The origin (center
    of the weaker flow) is indicated by a red star, the center of
    curvature of the apex of the bow shock is indicated by a green
    dot, and the geometric center of the conic section is indicated by
    a blue cross, which is offset by \(x_0\) from the origin.  Note
    that \(R_0\), \(R_{90}\), \(R_c\), \(a\), and \(b\) are all
    \emph{lengths} and are always positive, whereas \(x_0\) is a
    \emph{displacement} and may be positive or negative.}
  \label{fig:conics}
\end{figure*}
\begin{figure}
  \includegraphics[width=\linewidth]{figs/conic1}
  \caption{Example of a family of conic sections, all with the same
    planitude (flatness at apex, marked by white dot):
    \(\Pi = R_c/R_0 = 2 \). The quadric angle,
    \(\theta_q = \tan^{-1} T_q\), varies over the family (see text), with
    lower values of \(\theta_q\) giving larger values of the alatude,
    \(\Lambda = R_{90}/R_0 \), meaning more open wings.  Different values of
    \(\Pi\) can be achieved for the same quadrics by sliding them along
    the \(x\)-axis, which will also change the axis scales since these
    are normalized by \(R_0\).}
  \label{fig:conics-family}
\end{figure}

@@

@@

%Instead of the excentricity, we utilize the parameter $\theta_c$ to characterize the different curves, where
%$\tan\theta_c = \frac{b}{a}$,  $b$ and $a$ are the typical parameters of conics. A positive value for $\theta_c$ indicates that the given curve is a closed one, i.e
%an ellipse, while a negative value indicates that is an hyperbola. %Insert figures if neccesary

@@

A general parametric form\footnote{%
  The special case of the parabola needs to be treated differently,
  see Appendix~\ref{app:parabola}.} %
for the \(xy\) coordinates of the quadrics (in the \(\phi = 0\) plane,
and with the star at the origin) as a function of \(t = [0, \pi]\) is
then
\begin{gather}
  \label{eq:par-xy}
  \begin{aligned}
    x &= x_0 + \sigma a \Cos(t) \\ 
    y &= b\Sin(t) 
  \end{aligned}
\end{gather}
where
\begin{equation}
  \label{eq:sin-sinh-etc}
  \Sin{}, \Cos = \begin{cases}
    \sin{}, \cos & \text{ellipse}\\
    \sinh{}, \cosh & \text{hyperbola}
  \end{cases}
\end{equation}
Except for the circle case (\(\sigma = +1\), \(a = b\)), the parametric
variable \(t\) is not actually an angle in physical space.  Instead,
the polar form of the bow shape \(R(\theta)\) must be found by substituting
equations~\eqref{eq:par-xy} into \(\theta = \tan^{-1} y/x\) and
\(R = (x^2 + y^2)^{1/2}\).

The type of quadric surface can be characterized by the
\textit{quadric parameter}:
\begin{equation}
  \label{eq:Tq}
  T_q \equiv \sigma \frac{b} {a} \ , 
\end{equation}
where \(T_q < 0\) corresponds to open surfaces (hyperboloids) and
\(T_q > 0\) corresponds to closed surfaces (oblate spheroids with
\(T_q > 1\) and prolate spheroids with \(T_q < 1\)).  Special cases
are the sphere (\(T_q = 1\)) and the paraboloid (\(T_q = 0\)).
Alternatively, one can define a \textit{quadric angle}:
\(\theta_q = \tan^{-1} T_q\), which is marked in Figure~\ref{fig:conics}.
In the case of hyperboloids, the asymptotic opening angle of the wings
(\S~\ref{sec:plan-alat-bow} and Fig.~\ref{fig:characteristic-radii})
is \(\theta_\infty = \pi + \theta_q\) (note that \(\theta_q < 0\) in this case).

The set of parameters \((T_q, a, x_0)\) are then sufficient to
characterize the star/bow combination, where \(a\) is the quadric
scale and \(x_0\) is its center displacement from the star.  However,
we can also characterize the star/bow by \((R_0, \Pi, \Lambda)\), where
\(\Pi\) and \(\Lambda\) are the planitude and alatude, see
\S~\ref{sec:plan-alat-bow}.  We now derive the relationship between
these two descriptions.  The apex radius of curvature for a conic
section is
\begin{equation}
  \label{eq:Rc-conic}
  R_c = \frac{b^2}{a} \ ,
\end{equation}
whereas the perpendicular radius, \(R_{90}\), is the value of \(y\)
when \(x = 0\), which can be found from equations (\ref{eq:conic-x0},
\ref{eq:par-xy}) as
\begin{equation}
  \label{eq:conic-R90}
  R_{90} = \left|T_q\right|
  \left(2aR_0 - \sigma R_0^2\right)^{1/2} \ .
\end{equation}
Combining equations (\ref{eq:planitude}, \ref{eq:alatude},
\ref{eq:conic-x0}, \ref{eq:Tq}--\ref{eq:conic-R90}) yields
\begin{align}
  \label{eq:Tq-from-Pi-Lambda}
  T_q &= \sigma \left| 2 \Pi - \Lambda^2 \right|^{1/2} \\
  \label{eq:a-from-R0-Pi-Lambda}
  a &= R_0 \frac{ \Pi  } { \left|  2 \Pi - \Lambda^2\right| } \\
  x_0 &= R_0 \frac{\Pi - \Lambda^2} {2 \Pi - \Lambda^2}
\end{align}
From these we see that it is the sign of \(2 \Pi - \Lambda^2\) that determines
\(\sigma\) and thus whether a quadric is a spheroid or a hyperboloid.  For
example, for a constant planitude, \(\Pi\), we can have a family of
different quadric types, with varying alatude, \(\Lambda\), that increases
from oblate, through prolate and paraboloid, to hyperboloid, as
illustrated in Figure~\ref{fig:conics-family}.

@@

\begin{figure*}
  \centering
  \includegraphics[width=\linewidth]{figs/projected-R90-Rc-snapshots}
  \caption{Variation with inclination angle of projected quadric parameters}
  \label{fig:projected-R90-Rc-snapshots}
\end{figure*}
We now apply the machinery of \S~\ref{sec:projection} to find the projected shape of a quadric bow on the plane of the sky.

Once the parametrization is done, we can find the apparent shape of
the shell in the observer's frame, following the procedure explained
in section.  First of all, the intrinsic 3D shape of the shell is given by:
\begin{align}
x &= a\Cos(t)+x_0 \\ 
y &= b\Sin(t)\cos\phi \\
z &=  b\Sin(t)\sin\phi
\end{align}
The azimutal angle where the line of sight is tangent to the shell is given by equation (\ref{eq:tanphi}) and (\ref{eq:alpha}), then:
\begin{align}
\sin\phi_t &= \frac{b}{a}\tan i\Cot(t) 
\end{align}
where:
\begin{align}
\Cot(t) = \left\lbrace \begin{array}{c}
\cot t ~\mathrm{if~ellipse} \\
\coth t ~\mathrm{if~hyperbola}
\end{array}\right.
\end{align}
